%%---------------------------------------------------------------------------%%
%%
%% \file    draco/src/cdi_eospac/doc/cdi_eospac.tex
%% \author  Kelly Thompson
%% \date    Thu Apr 26 14:49:28 MST 2001
%% \brief   Using the EOSPAC Equation of State package plugin for CDI.
%%---------------------------------------------------------------------------%%
%% $Id$
%%---------------------------------------------------------------------------%%
\documentclass[11pt]{nmemo}
\usepackage[centertags]{amsmath}
\usepackage{amssymb,amsthm,graphicx}
\usepackage[mathcal]{euscript}
\usepackage{tabularx}
\usepackage{tmadd,tmath}
\usepackage{cite}
%\usepackage{c++}
\usepackage{fancycodes}  % needed to get ``LaTeX'' symbol.
\usepackage{float} % used for C++ example code.
\usepackage{shading}
\usepackage{fancybox}
\usepackage{url}
\usepackage[hang,small]{caption2}

%%---------------------------------------------------------------------------%%
%% DEFINE SPECIFIC ENVIRONMENTS HERE
%%---------------------------------------------------------------------------%%
%\newcommand{\elfit}{\ensuremath{\operatorname{Im}(-1/\epsilon(\vq,\omega)}}
%\msection{}-->section commands
%\tradem{}  -->add TM subscript to entry
%\ucatm{}   -->add trademark footnote about entry

\newcommand{\comp}[1]{\normalfont\normalsize\texttt{#1}}
\newcommand{\draco}{{\normalfont\sffamily Draco}}
\newcommand{\tycho}{{\normalfont\sffamily Tycho}}
\newcommand{\solon}{{\normalfont\sffamily Solon}}
\newcommand{\milagro}{{\normalfont\sffamily Milagro}}
\newcommand{\metis}{{\normalfont\sffamily Metis}}
\newcommand{\pcg}{{\normalfont\sffamily PCG}}
\newcommand{\cvs}{{\normalfont\bfseries CVS}}
\newcommand{\python}{{\normalfont\bfseries Python}}
\newcommand{\doxygen}{{\normalfont\bfseries Doxygen}}
\newcommand{\kcc}{{\normalfont\bfseries KCC}}
\newcommand{\xemacs}{{\normalfont\bfseries XEmacs}}
\newcommand{\bash}{{\normalfont\bfseries Bash}}

\floatstyle{plain}
\newfloat{cxxSampleCode}{!ht}{lom}
\floatname{cxxSampleCode}{Code Example}
\setlength{\belowcaptionskip}{5ex plus 4pt minus 2pt}

\newenvironment{doublequote}
{\begin{quote}\begin{quote}}
    {\end{quote}\end{quote}}

\newenvironment{codeExample}
{\footnotesize
  \VerbatimEnvironment
  \begin{SaveVerbatim}{\mycode}}%
  {\end{SaveVerbatim}%
  \noindent%
  \parashade[.950]{sharpcorners}{\gdef\outlineboxwidth{.5}%
    \UseVerbatim{\mycode}}\normalsize}

%%---------------------------------------------------------------------------%%
%% BEGIN DOCUMENT
%%---------------------------------------------------------------------------%%
\begin{document}

%%---------------------------------------------------------------------------%%
%% OPTIONS FOR NOTE
%%---------------------------------------------------------------------------%%

\toms{Distribution}
%\toms{Joe Sixpak/XTM, MS B226}
\refno{CCS-4:01-17(U)}
\subject{EOSPAC Equation of State Package for \draco}

%-------NO CHANGES
\divisionname{Computer and Computational Sciences Division}
\groupname{CCS-4: Transport Methods Group}
\fromms{Kelly Thompson CCS-4 D409}
\phone{(505) 665--3929}
\originator{kt}
\typist{kt}
\date{\today}
%-------NO CHANGES

%-------OPTIONS
%\reference{NPB Star Reimbursable Project}
%\thru{P. D. Soran, XTM, MS B226}
%\enc{list}
%\attachments{list}
%\cy{list}
%\encas
%\attachmentas
%\attachmentsas
%-------OPTIONS

%%---------------------------------------------------------------------------%%
%% DISTRIBUTION LIST
%%---------------------------------------------------------------------------%%

\distribution {
  CCS-4 MS D409:\\
  Group Leader:  J.E. Morel, X-6 MS D409\\
%  B.T. Adams, X-6 MS D409\\
%  R.E. Alcouffe, X-6 MS D409\\
%  M.L. Alme, X-6 MS D409\\
  D.G. Archuleta, X-6 MS D409\\
%  R.S. Baker, X-6 MS D409\\
%  F.W. Brinkley, Jr., X-6 MS D409\\
%  J.A. Dahl, X-6 MS D409\\
%  F.H. Dang, X-6 MS D409\\
  T.M. Evans, X-6 MS D409\\
%  S.A. Gallegos, X-6 MS D409\\
%  B.D. Ganapol, X-6 MS D409\\
  M.G. Gray, X-6 MS D409\\
%  J.C. Gulick, X-6 MS D409\\
%  M.L. Hall, X-6 MS D409\\
%  W.D. Hawkins, X-6 MS D409\\
  H.G. Hughes, X-6 MS D409\\
% M.A. Hunter, X-6 \\
 H. Lichtenstein, X-6 MS D409\\
% D.R. Marr, X-6 MS D409\\
  J.M. McGhee, X-6 MS D409\\
%  D. Mihalas, X-6 MS D409\\
%  M.L. Murillo, X-6 MS D409\\
  G.L. Olson, X-6 MS D409\\
  S. Pautz, X-6 MS D409\\
%  A.K. Prinja, X-6\\
%  E.W. Qarsen, X-6 MS D409\\
  R.M. Roberts, X-6 MS D409\\
  K.G. Thompson, X-6 MS D409\\
%  S.A. Turner, X-6 MS D409\\
%  W.T. Urban, X-6 MS D409\\
  T.J. Urbatsch, X-6 MS D409\\
%  T.A. Wareing, X-6 MS D409\\
  J.S. Warsa, X-6 MS D409\\
}

%%---------------------------------------------------------------------------%%
%% BEGIN NOTE
%%---------------------------------------------------------------------------%%

\opening

%%---------------------------------------------------------------------------%%
%%                           Introduction
%%---------------------------------------------------------------------------%%

\section{Introduction}
\label{intro}

The \texttt{cdi\_eospac} package has been introduced into the \draco\
project to provide a solution to two issues.  First, it provides a C++
object based interface to LANL's Sesame Tables (equation of state)
data.  These data tables have been produced by Thoeretical Division's
Equation of State and Mechanics of Materials Group (T-1,
\url{http://t1web.lanl.gov/}).  Second, it acts as a plug-in component
for \draco's Common Data Interface (CDI) package.  A list of four
components that make up the \texttt{cdi\_eospac} package are shown
below in Table~\ref{tab:components}.

This initial release of the \texttt{cdi\_eospac} package only provides
access to the following types of EOSPAC data:

\begin{itemize}
\item Temperature-based specific electron internal energy (kJ/g).
\item Temperature-based specific electron heat capacity (kJ/g/keV).
\item Temperature-based specific ion internal energy (kJ/g).
\item Temperature-based specific ion heat capacity (kJ/g/keV).
\item Average number of free electrons per ion.
\item Temperature-based specific electron thermal conductivity ($s^{-1}cm^{-1}$).
\end{itemize}

This list of available queries will expand to include more data types.
Additionally, this package has only been tested for the IRIX64 64-bit
architecture.  Versions for Linux, SunOS, Tru64 and 32-bit IRIX64 will
be released shortly.

% h - here, t - top, p - page
\begin{table}[!ht]
%  \begin{quote}
  \setcaptionwidth{0.8\textwidth}
  \caption{This package consists of three classes and a C++ wrapper for
    the EOSPAC vendor library.}
  \label{tab:components}
%  \end{quote}
  \footnotesize

    \begin{center}
      \newcolumntype{Y}{>{\raggedleft\setlength{\hsize}{0.3\linewidth}\arraybackslash}X}
      \newcolumntype{Z}{>{\raggedright\setlength{\hsize}{0.7\linewidth}\arraybackslash}X}
                                %    \begin{tabularx}{\linewidth}{YZ}
      \begin{tabularx}{0.8\linewidth}{YZ}
        \multicolumn{1}{c}{\textbf{Class}} & \multicolumn{1}{c}{\textbf{Role}} \\

        \hline
        \\

        \textbf{SesameTables}    & Provides a map between the available
                                   EOSPAC data types that may be
                                   returned by the library and numeric
                                   material identifiers. \\
        \\
        \textbf{Eospac}          & Provides access to equation of state
                                   data for a single material whose
                                   properties are defined by
                                   \bf{SesameTables}.\\
        \\
        \textbf{EospacWrapper}   & Provides C++ style access to the
                                   EOSPAC library functions ensuring
                                   correct data type translation. \\
        \\
        \textbf{EospacException} & Provides detailed information about
                                   and handle any exception thrown by
                                   the EOSPAC wrapper or library. \\
    \end{tabularx}
  \end{center}
  \normalsize
\end{table}

The \texttt{cdi\_eospac} package may be used as a stand-alone package;
however, it has been designed as a plug-in component for CDI.  Because
of this, its design follows the interface defined by the CDI package.
In particular, \texttt{cdi\_eospac} offers no mechanism for mixing
materials.  The only mixtures that it can utilize are those included
in the Sesame data tables provided by T-1 (e.g.: water).

This package provides a wrapper to data access libraries written,
supported and provided by X-division's data team
(EOSPAC~\cite{eospac-cranfill,eospac-web-site}).  Because this wrapper
interacts directly with the Sesame libraries it will only function on
machines that have these libraries available locally. The \draco\
project keeps a copy of the EOSPAC libraries for revision control
purposes.  The EOSPAC libraries are located at
\url{/radtran/vendors/eospac}.  The \draco\ project does not keep a
copy of the Sesame tables.


%%---------------------------------------------------------------------------%%
%%                             SesameTables
%%---------------------------------------------------------------------------%%

\section{SesameTables}
\label{SesameTables}

Each Eospac object provides a single point of access for extracting
(interpolating) equation of state data from the Sesame table files
located on X-division machines.  These files may contain lookup
tables that overlap each other (e.g.: different revisions of the same
data, see \url{http://int.lanl.gov/projects/sdm/win/materials/} for
specifics).  A SesameTables object provides a mechanism to explicitly
define the lookup tables that are to be used for retrieving equation
of state data.

A SesameTables object defines the \emph{single} material that an
Eospac object will allow access to.  It also specifies what state data
is required and tells Eospac to only cache the lookup tables for the
requested data types.

%----------------------------------------
% SesameTables - Instantiation
%----------------------------------------

\subsection{Instantiation Example}

In Code Example \ref{codeExample:SesameTablesInstantiation} below we
want to create a SesameTables object for Aluminum.  We initially
declare the SesameTables object and name it \texttt{AlSt}.  However,
it is not a useful object until we specify what state data we want to
be able to query and which specific Sesame tables to use for each
lookup.  The types of data that are available from EOSPAC are listed
in Table \ref{tab:SesameDataTypes} along with a brief description, the
associated table category index and the Sesame catalog number.  This
information is also available from the EOSPAC online manual
(\url{http://laurel.lanl.gov/XCI/PROJECTS/DATA/eos/UsersDocument/HTML/EOSPAC.html#5.4}),
and from the EOS/Sesame Data Information web site
(\url{http://int.lanl.gov/projects/sdm/win/materials/}).

\begin{cxxSampleCode}
\begin{codeExample}
#include "cdi_eospac/SesameTables.hh"

main()
{
  // Create a SesameTables object for Aluminum.
  // Must still assign material identifiers to one or more EOSPAC data types.

  rtt_cdi_eospac::SesameTables AlSt;

  // This matID for Al has lookup tables for enelc.
  // Al has id=371 and this is table=7 for data category=0.

  const int Al3717 = 3717;

  // This matId for Al has has lookup tables for zfree3.
  // Al has id=371, this is table=4 for data category=2.

  const int Al23714 = 23714;

 // Give the requested data types to AlSt and their associated material
 // identifiers (both for Al).

  AlSt.enelc( Al3717 ).zfree3( Al23714 );
}
\end{codeExample}
\caption{Example of instantiating a SesameTables object.}
\label{codeExample:SesameTablesInstantiation}
\end{cxxSampleCode}

\begin{table}[!ht]
  \setcaptionwidth{0.9\textwidth}
  \caption{EOSPAC provides access to 36 equation of state lookup
    tables.  These fall into three table categories.  These tables can
    also be queried for derivative information.}
  \label{tab:SesameDataTypes}
  \footnotesize

    \begin{center}
      \newcolumntype{C}{>{\centering\setlength{\hsize}{0.1\linewidth}\arraybackslash}X}
      \newcolumntype{D}{>{\centering\setlength{\hsize}{0.15\linewidth}\arraybackslash}X}
%      \newcolumntype{E}{>{\centering\setlength{\hsize}{0.1\linewidth}\arraybackslash}X}
%      \newcolumntype{R}{>{\raggedleft\setlength{\hsize}{0.15\linewidth}\arraybackslash}X}
      \newcolumntype{L}{>{\raggedright\setlength{\hsize}{0.6\linewidth}\arraybackslash}X}
      \begin{tabularx}{0.9\linewidth}{CDDL}
%        \multicolumn{1}{c}{\textbf{Data Type}} &
%        \multicolumn{1}{c}{\textbf{Sesame Catalog Number}} &
%        \multicolumn{1}{c}{\textbf{Table Category}} &
%        \multicolumn{1}{l}{\textbf{Description}} \\

        \textbf{Data Type} &
        \textbf{Sesame Catalog Number} &
        \textbf{Table Category} &
        \multicolumn{1}{c}{\textbf{Description}}

        \\
        \hline
        \\

        \textbf{prtot} & 301 & 0 & Temperature-based total pressure\\
        \textbf{entot} & 301 & 0 & Temperature-based total internal energy/mass\\
        \textbf{tptot} & 301 & 0 & Pressure-based total temperature\\
        \textbf{tntot} & 301 & 0 & Energy-based total temperature\\

        \textbf{pntot} & 301 & 0 & Energy-based total pressure\\
        \textbf{eptot} & 301 & 0 & Pressure-based total internal energy/mass\\

%            // Sesame Catalog 303 entries:

        \textbf{prion} & 303 & 0 & Temperature-based ion pressure\\
        \textbf{enion} & 303 & 0 & Temperature-based ion internal energy/mass\\
        \textbf{tpion} & 303 & 0 & pressure-based ion temperature\\
        \textbf{tnion} & 303 & 0 & energy-based ion temperature\\
        \textbf{pnion} & 303 & 0 & energy-based ion pressure\\
        \textbf{epion} & 303 & 0 & pressure-based ion internal energy/mass\\

%            // Sesame Catalog 304 entries:

        \textbf{prelc} & 304 & 0 & Temperature-based electron pressure\\
        \textbf{enelc} & 304 & 0 & Temperature-based electron internal energy/mass\\
        \textbf{tpelc} & 304 & 0 & Pressure-based electron temperature\\
        \textbf{tnelc} & 304 & 0 & Energy-based electron temperature\\
        \textbf{pnelc} & 304 & 0 & Energy-based electron pressure\\
        \textbf{epelc} & 304 & 0 & Pressure-based electron internal energy/mass\\

%            // Sesame Catalog 306 entries:

        \textbf{prcld} & 306 & 0 & Temperature-based cold curve pressure\\
        \textbf{encld} & 306 & 0 & Temperature-based cold curve internal energy/mass\\

%            // Sesame Catalogs 502-505

            \textbf{opacr}  & 502 & 1 & Temperature-based Rosseland mean opacity\\
            \textbf{opacc2} & 503 & 1 & Temperature-based electron conductive opacity\\
            \textbf{zfree2} & 504 & 1 & Temperature-based number of free electrons per ion\\
            \textbf{opacp}  & 505 & 1 & Temperature-based Planck mean opacity\\

%            // Sesame Catalogs 601-605

            \textbf{zfree3} & 601 & 2 & Temperature-based number of free electrons per ion\\
            \textbf{econde} & 602 & 2 & Temperature-based electron electrical conductivity\\
            \textbf{tconde} & 603 & 2 & Temperature-based electron thermal conductivity\\
            \textbf{therme} & 604 & 2 & Temperature-based electron thermo-electric coef\\
            \textbf{opacc3} & 605 & 2 & Temperature-based electron conductive opacity\\

%            // Sesame Catalog 411

            \textbf{tmelt} & 411 & 3 & Temperature-based melting temperature\\
            \textbf{pmelt} & 411 & 3 & Temperature-based melting pressure\\
            \textbf{emelt} & 411 & 3 & Temperature-based melting internal energy/mass\\

%            // Sesame Catalog 412

            \textbf{tfreez} & 412 & 3 & Temperature-based freezing temperature\\
            \textbf{pfreez} & 412 & 3 & Temperature-based freezing pressure\\
            \textbf{efreez} & 412 & 3 & Temperature-based freezing internal energy/mass\\

%            // Sesame Catalog 431

            \textbf{shearm} & 431 & 3 & Temperature-based shear modulus\\

    \end{tabularx}
  \end{center}
  \normalsize
\end{table}

Our code example below assumes that we only want to query for two
types of EOS data: temperature-based electron internal energies
(needed to obtain electron-based heat capacities, $C_{v_e} =
\frac{\partial E_e} {\partial T}$) and the average number of free
electrons per ion.  From Table \ref{tab:SesameDataTypes}, we see that
we need to request the data types ``enelc'' and ``zfree3''.  This
tells us what ``set'' functions to call to correctly configure
\texttt{AlSt}, but we must still assign material/table identifiers
(for Al) to each of these data types.

The table identifier for EOSPAC is an integer of the form:
[\emph{table-category}][\emph{material-index}][\emph{table-version}].
We can find these table identifiers by visiting
\url{http://int.lanl.gov/projects/sdm/win/materials/} and searching on
\emph{aluminum}.  This search brings up a list of available Sesame
tables.  We quickly see that the \emph{material-index} for aluminum is
371, and that there are several \emph{table-versions} available for
each \emph{table-category}. From Table \ref{tab:SesameDataTypes}, we
determine that we will need tables 304 (enelc) and 601 (zfree3).  In
the code example, we choose the table identifiers 3717 and 23714,
respectively, to meet our requirements.

The SesameTables object is fully defined by using the \emph{set}
functions for \texttt{enelc} and \texttt{zfree3}.  We can chain this
set functions together as shown in the code example.  Once the
SesameTables object has been configured it can be used in the
instantiation of an Eospac object (see the next section of this
documents for details).

%----------------------------------------
% SesameTables - General Use
%----------------------------------------

\subsection{General Use}

All of the data types listed in Table \ref{tab:SesameDataTypes} are
valid \emph{set} functions for SesameTables objects.  However, the
\texttt{cdi\_eospac} package only provides access routines for the
data types listed in \S\ref{intro} (\texttt{enelc}, \texttt{enion},
\texttt{zfree3}, and \texttt{tconde}).  As our needs expand this
package will provide access all of the data types that EOSPAC knows
about.

In some situations a SesameTables object will only be used once and so
an explicit instantiation may not be required.  The SesameTables
object can be a temporary that is passed as an argument to the Eospac
contructor.  An example of this usage is shown below in Code Example
\ref{codeExample:EospacInstantiation}.



%%---------------------------------------------------------------------------%%
%%                                Eospac
%%---------------------------------------------------------------------------%%

\section{Eospac}

The Eospac object provides access to \emph{Equation of State} data
found in the Sesame Tables (\texttt{sesame}, \texttt{sescu},
\texttt{sescu1}, \texttt{sescu9} and \texttt{sesou}) that are
available on the ACL (\url{/usr/projects/data}) and X-Division
(\url{/usr/local/codes/data}) networks.

An Eospac object represents a \emph{single} material that is defined
by a SesameTables object.  Please refer to the previous section
(\S\ref{SesameTables}) for details about SesameTables objects and
material identifiers.  To query an Eospac object the user must
provided a temperature in keV and a density in g/$cm^3$.  Alternately,
the user may provide equal length vectors of temperatures and
densities to the access routines.  The returned vector will contain
the interpolated values of the requested data type.

Eospac objects inherit their interface from the EOS interface
definitions declared in the cdi package (see cdi/EOS.hh).

In this release of \texttt{cdi\_eospapc}, no specialized smoothing of
the data will be preformed.  Additionally, all interpolation uses a
bi-linear scheme.  If needed, this package can be expanded to provide
some of the more advanced smoothing and interpolation schemes
available in EOSPAC.

%\begin{itemize}
%\item Interpolation scheme = bilinear
%\item Assumed default values?
%\end{itemize}

%----------------------------------------
% Eospac - Instantiation
%----------------------------------------

\begin{cxxSampleCode}
\begin{codeExample}
#inlcude "cdi_eospac/SesameTables.hh"
#inlcude "cdi_eospac/Eospac.hh"

main()
{
  // An Eospac object allows the user to access EoS information about a material that has been
  // constructed in a SesameTable object.  The constructor for Eospac takes one argument: a
  // SesameTables object.

  std::shared_ptr< rtt_cdi_eospac::Eospac > spAlEospac;

  // See the previous code example for a description of these two const values.

  const int Al3717  = 3717; const int Al23714 = 23714;

  // Try to instantiate the new Eospac object.  Simultaneously, we are assigned material IDs to
  // more SesameTable values.

  spAlEospac = new
    rtt_cdi_eospac::Eospac( rtt_cdi_eospac::SesameTables().enion( Al3717 ).tconde( Al23714 ) ) );
 }
\end{codeExample}
\caption{Example of instantiating an Eospac object.}
\label{codeExample:EospacInstantiation}
\end{cxxSampleCode}

\subsection{Instantiation}

The state of an Eospac object is determined by the SesameTables object
provided during instantiation.  The Eospac object takes the list of
requested data types and material identifiers and caches all of the
requested data tables.  The Eospac object will only be valid if the
associated SesameTables object defines at least one data type.

An example of instantiating an Eospac object is shown in Code Example
\ref{codeExample:EospacInstantiation}.  In this example, we declare a
smart pointer to an Eospac object and then intantiate the Eospac
object.  The single argument provided to the Eospac constructor is a
SesameTables object generated \emph{on-the-fly}.  The user can also
choose to create a presistent SesameTables object as demonstrated in
Code Example~\ref{codeExample:SesameTablesInstantiation}.  This object
represents the material Al (material 371) and has cached the data
tables required accessing information about thermal conductivity and
specific ion internal energy (and specific ion heat capacity, $C_{v_i} =
\frac{\partial E_i} {\partial T}$).


%----------------------------------------
% Eospac - Access Functions
%----------------------------------------

\subsection{Access Functions}

The Eospac class has been designed as a mechanism to obtain
interpolated equation of state data from the Sesame Table data files
(\texttt{sesame}, \texttt{sescu}, \texttt{sescu1}, \texttt{sescu9} and
\texttt{sesou}).  As such, Eospac has many access functions as shown
below in Table \ref{tab:accessFunctions}.  Examples demonstrating the
use of these accessor functions can be found in the next subsection
and in the unit tests ``cdi\_eospac/test/tEospac'' and
``cdi\_eospac/test/tEospacWithCDI.''

%Other functions are also
%available that allow the user to query the opacity lookup table and to
%obtain text descriptions of the data represented by this object.

Some of the more commonly used member functions for Eospac are listed
in Table \ref{tab:accessFunctions} below.  These access functions
expect temperatures to be given in keV and densities to be given in
g/cm$^3$.  The interpolated equation of state values returned by these
member functions will have the units specified in the table.

Eospac provides an associated set of vector access functions.  These
member functions have the same name as those shown in
Table \ref{tab:accessFunctions}, but the temperature and density
arguments have type \texttt{const~vector< double >\&} and the return
type is a \texttt{vector< double >}.


\begin{table}[H]%[!ht]
    \caption{Eospac scalar member functions.}
    \label{tab:accessFunctions}
    \footnotesize

    \begin{center}
      \newcolumntype{Y}{>{\raggedright\setlength{\hsize}{0.5\linewidth}\arraybackslash}X}
      \newcolumntype{Z}{>{\raggedright\setlength{\hsize}{0.5\linewidth}\arraybackslash}X}
      \begin{tabularx}{0.9\linewidth}{YZ}
        \multicolumn{1}{c}{\textbf{Eospac Member Functions}} & \multicolumn{1}{c}{\textbf{Role}} \\

        \hline
        \\
\texttt{double~getSpecificElectronInternalEnergy(}
\texttt{\mbox{ double~temp,}}
\texttt{\mbox{ double~density ) const}}
& Given the temperature (keV) and density(g/$cm^3$) information, return a bi-linearly
  interpolated specific electron internal energy value (kJ/g).  \\
&  Requires \texttt{enelc}. \\
\\
\texttt{double~getElectronHeatCapacity(}
\texttt{\mbox{ double~temp,}}
\texttt{\mbox{ double~density)}}
& Given the temperature (keV) and density(g/$cm^3$) information, return a bi-linearly
  interpolated specific electron heat capacity value (kJ/g/keV).\\
&  Requires \texttt{enelc}.\\
\\
\texttt{double~getSpecificIonInternalEnergy(}
\texttt{\mbox{ double~temp,}}
\texttt{\mbox{ double~density ) const}}
& Given the temperature (keV) and density(g/$cm^3$) information, return a bi-linearly
  interpolated specific ion internal energy value (kJ/g).  \\
&  Requires \texttt{enion}. \\
\\
\texttt{double~getIonHeatCapacity(}
\texttt{\mbox{ double~temp,}}
\texttt{\mbox{ double~density)}}
& Given the temperature (keV) and density(g/$cm^3$) information, return a bi-linearly
  interpolated specific ion heat capacity value (kJ/g/keV).\\
&  Requires \texttt{enion}.\\
\\
\texttt{double~getNumFreeElectronsPerIon(}
\texttt{\mbox{ double~temp,}}
\texttt{\mbox{ double~density ) const}}
& Given the temperature (keV) and density(g/$cm^3$) information, return a bi-linearly
  interpolated average number of electrons per ion.  \\
&  Requires \texttt{zfree3}. \\
\\
\texttt{double~getElectronThermalConductivity(}
\texttt{\mbox{ double~temp,}}
\texttt{\mbox{ double~density)}}
& Given the temperature (keV) and density(g/$cm^3$) information, return a bi-linearly
  interpolated electron thermal conductivity ($cm^{-1}s^{-1}$).\\
&  Requires \texttt{tconde}.\\
\\
      \end{tabularx}
    \end{center}
    \normalsize
\end{table}

%In Addition to these \emph{standard} access functions, both
%GandolfOpacity objects have STL-like accessor routines.  These
%accessors can be used if the desired opacity container is not a
%\texttt{vector<double>}.  The opacity container must be a 1-D
%container for both GandolfGrayOpacity and GandolfMultigroupOpacity.

An example using Eospac to obtain electron heat capacity and thermal
conductivity values for Aluminum is shown blow in Code Example~%
\ref{codeExample:eospacUse_part1} and
\ref{codeExample:eospacUse_part2}.  For more examples please examine
the \texttt{tEospac.cc} and \texttt{tEospacWithCDI.cc} unit tests.

\begin{cxxSampleCode}
\begin{codeExample}
#include "cdi_eospac/Eospac.hh"
#include "cdi_eospac/SesameTables.hh"
#include "cdi/CDI.hh"

main()
{
  // Material identifiers for Al (see previous code examples for details).

  const int Al3717 = 3717;
  const int Al23714 = 23714;

  // Declare a smart pointer to Eospac.  Eospac is derived from the interface defined by
  // rtt_cdi:EoS.

   std::shared_ptr< rtt_cdi::EoS > spEosAl;

   // Use a try block when instantiating the Eospac object so that we can catch any thrown
   // exceptions.

   try
      {

        // Create the SesameTables object within the instantiation request.  Request enelc
        // (electron heat capacity) and tconde (electron thermal conductivity) data for Al
        // to be cached.

         spEosAl = new
            rtt_cdi_eospac::Eospac(
               SesameTables().enelc( Al3717 ).tconde( Al23714 ) );
      }
   catch ( const rtt_cdi_eospac:EospacException& EosError )
      {
         // Instantiation failed so print a report and exit.
         cout << EosError.what();
         return -1;
      }

   // Set some input parameters.

   double temp = 0.0005; // keV
   double dens = 1.15;   // g/cm^3



   // continued below ...

\end{codeExample}

\caption{Example of using a Eospac object (part 1 of 2)}
\label{codeExample:eospacUse_part1}
\end{cxxSampleCode}

\begin{cxxSampleCode}
\begin{codeExample}
   // continued from above ...

   // Interpolate and return a heat capacity value (kJ/g/keV) for Al at (temp, dens).

   double Cve = spEosAl->getElectronHeatCapacity( temp, dens );

   // Interpolate and return an a vector of electron thermal conductivity values for Al for each
   // pair of temperature (keV) and density (g/cm^3) values.

   const int ndata = 3;
   vector< double > vtemp( ndata );
   vector< double > vdens( ndata );
   for ( int i=0; it<ndata; ++i)
   {
      vtemp[i] = 0.0005*(i+1);  // keV
      vdens[i] = 1.0*(i+1);     // g/cm^3
   }

   vector< double > vThermCond = spEosAl->getElectronThermalConductivity( vtemp, vdens );

   return 0;
}
\end{codeExample}
\caption{Example of using a Eospac object (part 2 of 2).}
\label{codeExample:eospacUse_part2}
\end{cxxSampleCode}


%%---------------------------------------------------------------------------%%
%%                  EospacException and EospacWrapper
%%---------------------------------------------------------------------------%%

\pagebreak

\section{EospacException and EospacWrapper}

The EospacWrapper routines provide access to the FORTRAN EOSPAC
libraries.  This set of routines act as the interface between the
FORTRAN library and the C++ objects that make up the
\texttt{cdi\_eospac} package.  A client code could theoretically use
the functions found in EospacWrapper to access the FORTRAN Gandolf
libraries directly without using the rest of the package.  However,
the design of this package did not have this approach in mind.

The client code may have some indirect interaction with the
EospacException object if the client is interested in intercepting
exceptions thrown by various routines in this package.  The client
code should include a try block similar to the one shown in Code
Example~\ref{codeExample:EospacInstantiation} when instantiating an
Eospac object.  There is no requirement for the client code to
directly interact with either of these classes.


%%---------------------------------------------------------------------------%%
%%                         EOSPAC Libraries
%%---------------------------------------------------------------------------%%

%\newpage

\section{EOSPAC Libraries}

Information about the EOSPAC libraries can be found online at
\url{http://laurel.lanl.gov/PROJECTS/DATA/eos/}.  This web site contains
information about the location of the EOSPAC libraries and Sesame libraries
alongside the online EOSPAC manual.  Links to information on the Sesame
libraries and material identifiers also exist on this web site.

%%---------------------------------------------------------------------------%%

%\pagebreak
\bibliographystyle{../../../../../draco/doc/tex/rnote}
\bibliography{../../../../../draco/doc/bib/draco}

\closing
\end{document}

%%---------------------------------------------------------------------------%%
%% end of cdi_eospac.tex
%%---------------------------------------------------------------------------%%
